---
title: "Homework 2"
author: Natalie Boychuk 
date: September 27 2020 
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```


## Problem 1 

Read the Mr. Trashwheel dataset. 

```{r}
trashwheel_df = 
	read_xlsx(
		path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)

```

Read precipitation data for 2017/2018! 

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year) 

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year) 

```

Combine annual precipitation 

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name 
  ) 

precip_df = 
  bind_rows(precip_2018, precip_2017) %>%  view

precip_df =
	left_join(precip_df, month_df, by = "month") 

```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches and the median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.


## Problem 2 

```{r reading in dataset and doing some basic cleaning}
subway_df = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") 
  subway_df = janitor::clean_names(subway_df) %>%  
  select(line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entrance_type, entry, vending, ada) %>% 
    mutate(entry = recode(entry, YES = "TRUE", NO = "FALSE")) %>% 
    mutate(entry = as.logical(entry))
 
  
```


```{r reformatting df to separate route number and route type}
clean_subway_df = subway_df %>% 
  mutate(route8 = as.character(route8),
         route9 = as.character(route9),
         route10 = as.character(route10),
         route11 = as.character(route11)
         ) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_no",
    values_to = "route_name",
    names_prefix = "route"
    ) 

view(clean_subway_df)

clean_subway_df %>% 
  filter(ada == TRUE) %>% 
  distinct(station_name, line)

```

```{r}
distinct(clean_subway_df, station_name, line)
```


This dataset provides data on transit 

There are a total of 455 distinct subway stations in NYC. Of those, 84 distinct stations are ADA compliant. 

```{r}
clean_subway_df %>% 
  filter(entry == "TRUE",
         vending == "NO")
```





